Memes: an R interface to the MEME Suite

### 2020-10-27 manuscript.Rmd
typo fix

 - as mentioned previously, the multidimensional naature of meme results is difficult to represent
 - Output data usually contain motifs & descriptive statistics about those motifs
 - motif info can be stored in universalmotif format, however, not enough space for data
 - data.frames do good job of storing data (incl nested data), but cannot store motifs
 - we solve this problem by storing the unviversalmotif inside the data.frame (which is valid R)
 - we add nested list columns as needed to link multidimensional data
 - advantage of this approach over an S4 is all normal tools work on it (tidyverse), joins, etc.

Many of the MEME Suite tools return files contianing descriptive
properties of motifs. For example, MEME and DREME return de-novo discovered
motifs from query sequences along with statistical information about their
enrichment. The `universalmotif` data structure is a powerful R/Bioconductor
representation for motif matrices and their associated metadata
\cite{universalmotif}.  However, `universalmotif` objects are not ideal for
performing bulk manipulation or queries of metadata, nor do they support complex
data structures such as nested elements. Many MEME Suite analysis pipelines
involve identifying motifs which are further analyzed by multiple tools,
creating additional metadata which cannot be easily tracked in bulk using
`universalmotif` format. Thus an ideal representation for these data is one
which can hold an unlimited amount of metadata, contain complex hierarchies, and
be easily manipulated using standard analysis tools. `memes` therefore extends
the `universalmotif` format by providing data.frames holding motif metadata
which retain links to their `universalmotif` representation. These
universalmotif data.frames combine the advantages of the `universalmotif`
framework for motif manipulation with the extensibility and flexibility of R
data.frames. These `universalmotif` data.frames form the basis for a majority of
`memes` data outputs. These structures are also valid input types to `memes`
functions, and when used as such, output data are appended as new columns to the
input data, ensuring data provenance.

>>>>

# CONTINUE
 - as mentioned previously, the multidimensional naature of meme results is difficult to represent
 - Output data usually contain motifs & descriptive statistics about those motifs
 - motif info can be stored in universalmotif format, however, not enough space for data
 - data.frames do good job of storing data (incl nested data), but cannot store motifs
 - we solve this problem by storing the unviversalmotif inside the data.frame (which is valid R)
 - we add nested list columns as needed to link multidimensional data
 - advantage of this approach over an S4 is all normal tools work on it (tidyverse), joins, etc.

Many of the MEME Suite tools return files containing descriptive
properties of motifs. For example, MEME and DREME return de-novo discovered
motifs from query sequences along with statistical information about their
enrichment. The `universalmotif` data structure is a powerful R/Bioconductor
representation for motif matrices and their associated metadata
\cite{universalmotif}.  However, `universalmotif` objects are not ideal for
performing bulk manipulation or queries of metadata, nor do they support complex
data structures such as nested elements. Many MEME Suite analysis pipelines
involve identifying motifs which are further analyzed by multiple tools,
creating additional metadata which cannot be easily tracked in bulk using
`universalmotif` format. Thus an ideal representation for these data is one
which can hold an unlimited amount of metadata, contain complex hierarchies, and
be easily manipulated using standard analysis tools. `memes` therefore extends
the `universalmotif` format by providing data.frames holding motif metadata
which retain links to their `universalmotif` representation. These
universalmotif data.frames combine the advantages of the `universalmotif`
framework for motif manipulation with the extensibility and flexibility of R
data.frames. These `universalmotif` data.frames form the basis for a majority of
`memes` data outputs. These structures are also valid input types to `memes`
functions, and when used as such, output data are appended as new columns to the
input data, ensuring data provenance.
